Generating Random Logic Programs Using Constraint Programming

### 2020-01-16 paper/paper.tex
Separate the model from other functionality
Make constraint variables more readable
A redundant constraint for occurrences
Some redundant constraints

Parameters:
\begin{itemize}
\item a list of predicates $\predicates{}$,
\item a list of their arities $\arities{}$ (including zero),
  \begin{itemize}
  \item maximum arity $\maxArity{} \coloneqq \max \arities{}$.
  \end{itemize}
\item a list of variables $\variables{}$,
\item and a list of constants $\constants{}$.
  \begin{itemize}
  \item Each of them can be empty, but $|\constants{}| + |\variables{}| > 0$.
  \end{itemize}
\item a list of probabilities that are randomly assigned to clauses,
\item option to forbid all cycles or just negative cycles,
\item $\maxNumNodes{} \ge 1$: maximum number of nodes in the tree representation
  of a clause,
\item $\maxNumClauses{} \ge |\predicates{}|$: maximum number of clauses in a
  program,
\item maximum number of solutions,
\end{itemize}

@@

\section{Eliminating Variable Symmetries} \label{sec:variable_symmetry}

@@

\begin{definition}
  We define $\arrayd{introductions}{|\variables{}|}{\{ 0, \dots, N \}}$ such
  that for $v \in \variables{}$,
  \[
    \variable{introductions}[v] = \begin{cases}
      \min \variable{occurrences}[v] & \text{if }
      \variable{occurrences}[v] \ne \emptyset\\
      N & \text{otherwise.}
    \end{cases}
  \]
\end{definition}

@@

@@

\section{Entailment Checking for Negative/All Cycles}
% TODO: explain that the hasNegativeCycles algorithm is an extension of a simple
% cycle detection algorithm.

@@

\begin{itemize}
\item A constraint for logical equivalence. An algorithm to reduce each tree to
  some kind of normal form. Not doing this on purpose. Leaving for further work.
\item Perhaps negative cycle detection could use the same graph as the
  independence propagator? If we extend each domain to {-1, 0, 1}, but that
  might make propagation weaker or slower.
\item Could investigate how uniform the generated distribution of programs is.
  Distributions of individual parameters will often favour larger values
  because, e.g., there are more 5-tuples than 4-tuples.
\item Inference options to explore. Logspace vs normal space. Symbolic vs
  non-symbolic. Propagate evidence (might be irrelevant)? Propagate weights?
  Supported knowledge compilation techniques: sdd, sddx, bdd, nnf, ddnnf, kbest,
  fsdd, fbdd.
\item Mention the random heuristic. Mention that restarting gives better
  randomness, but duplicates become possible. Restarting after each run is
  expensive. Periodic restarts could be an option.
\end{itemize}

>>>>

Parameters:
\begin{itemize}
\item a list of predicates $\predicates{}$,
\item a list of their arities $\arities{}$ (including zero, but assuming that at
  least one predicate has non-zero arity),
  \begin{itemize}
  \item maximum arity $\maxArity{} \coloneqq \max \arities{}$.
  \end{itemize}
\item a list of variables $\variables{}$,
\item and a list of constants $\constants{}$.
  \begin{itemize}
  \item Each of them can be empty, but $|\constants{}| + |\variables{}| > 0$.
  \end{itemize}
\item a list of probabilities that are randomly assigned to clauses,
\item option to forbid all cycles or just negative cycles,
\item $\maxNumNodes{} \ge 1$: maximum number of nodes in the tree representation
  of a clause,
\item $\maxNumClauses{} \ge |\predicates{}|$: maximum number of clauses in a
  program,
\item maximum number of solutions,
\end{itemize}

@@

\section{Eliminating Variable Symmetries} \label{sec:variable_symmetry}
% TODO: only if we have multiple variables

@@

\begin{constraint}[Redundant]
  For $u \ne v \in \constants{} \cup \variables{} \cup \{ \Box \}$,
  \[
    \variable{occurrences}[u] \cap \variable{occurrences}[v] = \emptyset.
  \]
\end{constraint}

\begin{definition}
  We define $\arrayd{introductions}{|\variables{}|}{\{ 0, \dots, N \}}$ such
  that for $v \in \variables{}$,
  \[
    \variable{introductions}[v] = \begin{cases}
      1 + \min \variable{occurrences}[v] & \text{if }
      \variable{occurrences}[v] \ne \emptyset\\
      0 & \text{otherwise.}
    \end{cases}
  \]
\end{definition}

\begin{constraint}[Redundant]
  $\variable{allDifferentExcept0}(\variable{introductions})$.
\end{constraint}

@@

\subsection{A Redundant Constraint}

For each clause...

\begin{definition}
  We define an auxiliary set variable
  \[
    \variable{potentialIntroductions} \subseteq \{ 0, \dots, N \}
  \]
  by: for $i = 0, \dots, \maxNumNodes{} - 1$, if
  $\variable{treeValues}[i].\variable{name} \in \tokens{}$, then
  \[
    \variable{potentialIntroductions} \cap \{ \maxArity{} \times (i + 1) + j + 1
    \mid j = 0, \dots, \maxArity{} - 1 \} = \emptyset.
  \]
\end{definition} % TODO: maybe add to the example
In other words, if the node not a predicate, a variable cannot be introduced as
one of its `arguments'.

\begin{constraint}
  For $v \in \variables{}$, $\variable{introductions}[v] \in
  \variable{potentialIntroductions}$.
\end{constraint}

@@

\section{Entailment Checking for Negative/All Cycles}
% TODO: explain that the hasNegativeCycles algorithm is an extension of a simple
% cycle detection algorithm.
% TODO: the difficulty with propagation is that there is no good way to define
% the adjacency matrix with 0/-1/+1 values, the negativity condition is too
% difficult.
% TODO: NB: the set of decision variables for this constraint must exclude all
% arguments.

@@

\begin{itemize}
\item A constraint for logical equivalence. An algorithm to reduce each tree to
  some kind of normal form. Not doing this on purpose. Leaving for further work.
\item Perhaps negative cycle detection could use the same graph as the
  independence propagator? If we extend each domain to {-1, 0, 1}, but that
  might make propagation weaker or slower.
\item Could investigate how uniform the generated distribution of programs is.
  Distributions of individual parameters will often favour larger values
  because, e.g., there are more 5-tuples than 4-tuples.
\item Inference options to explore. Logspace vs normal space. Symbolic vs
  non-symbolic. Propagate evidence (might be irrelevant)? Propagate weights?
  Supported knowledge compilation techniques: sdd, sddx, bdd, nnf, ddnnf, kbest,
  fsdd, fbdd.
\item Mention the random heuristic. Mention that restarting gives better
  randomness, but duplicates become possible. Restarting after each run is
  expensive. Periodic restarts could be an option.
\item Could add statistics about what constraints tend to conflict.
\end{itemize}
